---
date: "2019-05-05T00:00:00+01:00"
title: "Importing data"
linktitle: "1 Importing data"
author: admin
type: book
weight: 1
commentable: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = TRUE,
  out.width = "100%"
)
knitr::opts_knit$set(root.dir = here::here("content/docs"))
```


## Loading included data

You can load the demo data in `microbiomedataset` package to see the `microbiome_dataset` class.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
library(microbiomedataset)
library(tidyverse)
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
data("global_patterns")
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
global_patterns
```

So you can see that we have 1,9216 variables and 26 samples in the dataset.

## Create microbiome_dataset class object

You can also create the `microbiome_dataset` class using the `create_microbiome_dataset` function. 

We need to prepare at least three data for it.

1. `expression_data`: rows are variables and columns are samples.

2. `sample_info`: Information for all the samples in `expression_data`. The first column should be `sample_id` which should be identical with the column names of `expression_data`.

3. `variable_info`: Information for all the variables in `expression_data`. The first column should be `variable_id` which should be identical with the row names of `expression_data`.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
expression_data <-
  as.data.frame(matrix(
    sample(1:100, 100, replace = TRUE),
    nrow = 10,
    ncol = 10
  ))

rownames(expression_data) <- paste0("OTU", 1:nrow(expression_data))
colnames(expression_data) <-
  paste0("Sample", 1:ncol(expression_data))

expression_data

variable_info <-
  as.data.frame(matrix(
    sample(letters, 70, replace = TRUE),
    nrow = nrow(expression_data),
    ncol = 7
  ))

rownames(variable_info) <- rownames(expression_data)
colnames(variable_info) <-
  c("Domain",
    "Phylum",
    "Class",
    "Order",
    "Family",
    "Genus",
    "Species")

variable_info$variable_id <-
  rownames(expression_data)

variable_info <-
  variable_info %>% 
  dplyr::select(variable_id, dplyr::everything())

sample_info <-
  data.frame(sample_id = colnames(expression_data),
             class = "Subject")

object <-
  create_microbiome_dataset(
    expression_data = expression_data,
    sample_info = sample_info,
    variable_info = variable_info
  )

object
```

## Convert phyloseq class to microbiome_dataset class object

We can also transfer or convert other common class object to `microbiome_dataset` class.

Please install `phyloseq` package at first.


```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
if(!require(BiocManager)){
  install.packages("BiocManager")
}

if(!require(phyloseq)){
  BiocManager::install("phyloseq")
}
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
library(phyloseq)
data(GlobalPatterns)
GlobalPatterns
```

The first function is `convert2microbiome_dataset`:

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
object1 <- 
  convert2microbiome_dataset(object = GlobalPatterns)
object1
```

The second function is `as.microbiome_dataset`:

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
object2 <- 
  as.microbiome_dataset(object = GlobalPatterns)
object2
```


```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
microbiomedataset::plot_barplot(object = object, 
                                top_n = 5,
                                fill = "Phylum")
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
microbiomedataset::plot_barplot(object = object, 
                                top_n = 5,
                                fill = "Phylum", 
                                relative = TRUE, 
                                re_calculate_relative = TRUE)
```

## Session information

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
sessionInfo()
```
