---
date: "2019-05-05T00:00:00+01:00"
title: "Functions for accessing and preprocessing data"
linktitle: "3 Functions for accessing and preprocessing data"
author: admin
type: book
weight: 3
commentable: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = TRUE,
  out.width = "100%"
)
knitr::opts_knit$set(root.dir = here::here("content/docs"))
```


## Loading included data

Let's load the demo data.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
library(microbiomedataset)
library(tidyverse)
data("global_patterns")
global_patterns
```

## Accessors

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
dim(global_patterns)
nrow(global_patterns)
ncol(global_patterns)
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
colnames(global_patterns)
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
head(rownames(global_patterns))
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
extract_sample_info(global_patterns) %>% 
  colnames()
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
extract_variable_info(global_patterns) %>% 
  colnames()
```


```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
extract_expression_data(global_patterns) %>% 
  head()
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
extract_sample_info(global_patterns) %>% 
  head()
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
extract_variable_info(global_patterns) %>% 
  head()
```

## Preprocessing

The `microbiomedataset` package also includes functions for filtering, subsetting, and merging abundance data.

In the following example, the `global_patterns` data is first transformed to relative abundance, creating the new `global_patterns2` object, which is then filtered such that only OTUs with a mean greater than 10^-5 are kept.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
global_patterns2 <-
  global_patterns %>%
  transform2relative_intensity() %>%
  mutate2variable(what = "mean_intensity") %>%
  activate_microbiome_dataset(what = "variable_info") %>%
  filter(mean_intensity > 10 ^ (-5))
```

This results in a highly-subsetted object, `global_patterns2`, containing just 4624 of the original ~19216 OTUs.

Next, only remain the variables that phylum Chlamydiae.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
global_patterns_chl <-
  global_patterns %>%
  activate_microbiome_dataset(what = "variable_info") %>%
  dplyr::filter(Phylum == "Chlamydiae")
```

Next, only remain the samples with total intensity > 20.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
global_patterns_chl <-
  global_patterns_chl %>%
  mutate2sample(what = "sum_intensity") %>%
  activate_microbiome_dataset(what = "sample_info") %>%
  filter(sum_intensity > 20)
```

## Session information

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
sessionInfo()
```
